% ===========================================================
%  hea-bench: an open benchmark suite and reference baselines
%  for high-entropy alloy phase prediction
%
%  Manuscript draft for SoftwareX (Elsevier).
%  Build with pdfLaTeX + biber:
%      pdflatex main && biber main && pdflatex main && pdflatex main
% ===========================================================

\documentclass[review,times,nonatbib]{elsarticle}

% Page geometry: one-inch margins, manuscript convention.
\usepackage[margin=1in]{geometry}

% Encoding and fonts. pdfLaTeX-compatible only.
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{textcomp}
\usepackage{textgreek}

% Math and units.
\usepackage{amsmath, amssymb}
\usepackage{siunitx}
\sisetup{detect-all, range-phrase = --, range-units = single}

% Tables and figures.
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{tabularx}

% Bibliography in APA style, numbered in text.
\usepackage[backend=biber,
            style=apa,
            citestyle=numeric-comp,
            sorting=none]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{refs.bib}

% Hyperlinks, set last per package-loading order conventions.
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{xspace}

% Lightweight macros for the few mathematical symbols that recur.
\newcommand{\Smix}{\ensuremath{\Delta S_{\mathrm{mix}}}}
\newcommand{\Hmix}{\ensuremath{\Delta H_{\mathrm{mix}}}}
\newcommand{\Tm}{\ensuremath{T_{\mathrm{m}}}}
\newcommand{\Omegaparam}{\ensuremath{\Omega}}
\newcommand{\hea}{\textsc{hea-bench}\xspace}

\journal{SoftwareX}

\begin{document}

\begin{frontmatter}

\title{%
  An open benchmark suite and reference baselines for high-entropy
  alloy phase prediction
}

\author[1]{David Fieser\corref{cor1}}
\ead{davjfies@gmail.com}
\cortext[cor1]{Corresponding author.}

\address[1]{Department of Mechanical and Aerospace, University of
Tennessee Knoxville, 1512 Middle Drive, Knoxville, TN 37996, USA.
ORCID \href{https://orcid.org/0009-0007-5754-4331}{0009-0007-5754-4331}}

\begin{abstract}
Empirical phase-prediction rules for high-entropy alloys are widely
used as fast pre-screens for candidate compositions. The four most
common rules are the entropy criterion of Yeh and colleagues
\cite{yeh2004}, the atomic-size mismatch criterion of Zhang and
colleagues \cite{zhang2008}, the valence-electron-concentration
criterion of Guo and Liu \cite{guo2011}, and the
\Omegaparam{}-parameter criterion of Yang and Zhang \cite{yang2012}.
The accuracies reported in the original publications, however, were
obtained on heterogeneous and mutually incompatible datasets, and
no published artifact treats the four rules as diagnostic
classifiers on a common open benchmark. This work introduces
\hea{}, an open Python library and reproducible benchmark suite
that consolidates three primary open phase-label datasets into a
single \num{7784}-composition resource with explicit per-row source
provenance and cross-source label-conflict tracking, implements
the four rules as binary or multi-class diagnostic classifiers with
full statistics (accuracy, sensitivity, specificity, Wilson 95\%
confidence intervals, receiver-operating-characteristic curves, and
Youden's $J$), and ships in two equivalent user surfaces. The first
is a standard installable Python package. The second is an
in-browser frontend that runs the same release wheel through
Pyodide on WebAssembly, requiring no local installation. On the
consolidated benchmark, both binary rules collapse to a
``predict-single-phase-almost-always'' regime with Youden's $J$
values of 0.075 and 0.032, and a sweep of the
\citeauthor{zhang2008}~\cite{zhang2008} $\delta$ threshold places
the dataset-optimal balance point at $\delta < \SI{2.5}{\percent}$
rather than the canonical $\delta < \SI{6.5}{\percent}$. The
package is released as version 0.0.1 under the MIT License with
more than 150 tests, including pinned regression tests on every
reported benchmark statistic.
\end{abstract}

\begin{keyword}
high-entropy alloys \sep
multi-principal element alloys \sep
phase prediction \sep
benchmark suite \sep
empirical descriptors \sep
Miedema model \sep
open data \sep
diagnostic classifiers
\end{keyword}

\end{frontmatter}

% ====================================================================
\section*{Required Metadata}
% ====================================================================

\begin{table}[h]
\caption{Code metadata.}
\centering
\begin{tabular}{p{0.39\linewidth} p{0.55\linewidth}}
\toprule
\textbf{Nr.} & \textbf{Code metadata description} \\
\midrule
C1 Current code version & 0.0.1 \\
C2 Permanent link to code/repository & To be assigned via Zenodo on
  the first tagged release \\
C3 Permanent link to reproducible capsule & Not applicable. The
  package is dependency-free in its core and reproducible via a
  standard package-manager install. \\
C4 Legal code license & MIT \\
C5 Code versioning system used & git \\
C6 Software code languages, tools, and services used & Python
  ($\geq$3.10), Pyodide for browser deployment \\
C7 Compilation requirements, operating environments, and dependencies
  & Pure Python, dependency-free core. Optional extras for tabular
  data handling, plotting, and testing. \\
C8 If available, link to developer documentation or manual &
  Repository readme file \\
C9 Support email for questions & \texttt{davjfies@gmail.com} \\
\bottomrule
\end{tabular}
\end{table}

% ====================================================================
\section{Motivation and significance}
% ====================================================================

High-entropy alloys, introduced independently by Cantor and
colleagues \cite{cantor2004} and by Yeh and colleagues
\cite{yeh2004} in 2004, are multi-component alloys with five or
more principal elements near equiatomic composition. The
configurational mixing entropy in such alloys is large enough to
stabilize disordered solid solutions over the intermetallic
compounds predicted by classical Hume--Rothery rules. The
compositional space is combinatorially larger than that of
conventional alloys, which makes pre-screening a practical
necessity before any expensive density-functional-theory, CALPHAD,
or experimental campaign. The practical importance of robust
pre-screening is illustrated by recent work on the thermodynamic
stability of femtosecond laser-ablated
CuCoMn$_{1.75}$NiFe$_{0.25}$ nanoparticles
\cite{fieser2025surface}, in which the choice of phase-prediction
descriptor materially affected the interpretation of stability
during the surface restructuring process.

Four empirical phase-prediction rules dominate the high-entropy
alloy literature. The first is the configurational mixing entropy
criterion of Yeh and colleagues \cite{yeh2004}, which classifies
an alloy as high-entropy when $\Smix > 1.5 R$. The second is the
atomic-size mismatch criterion of Zhang and colleagues
\cite{zhang2008}, which predicts a single-phase solid solution
when $\delta < \SI{6.5}{\percent}$. The third is the
valence-electron-concentration criterion of Guo and Liu
\cite{guo2011}, which predicts a face-centred-cubic structure at
$\mathrm{VEC} \geq 8.0$ and a body-centred-cubic structure at
$\mathrm{VEC} < 6.87$. The fourth is the
\Omegaparam{}-parameter criterion of Yang and Zhang
\cite{yang2012}, with $\Omegaparam = \Tm \, \Smix \, / \,
|\Hmix|$ and a threshold of $\Omegaparam > 1.1$ for single-phase
stability. Each rule was supported in its original publication by
classification accuracies near \SI{80}{\percent} on the dataset
available at the time.

Three gaps in the open literature motivate the present work.
First, every published evaluation of these rules uses a different
dataset, with different preprocessing, deduplication, and phase
labelling. Head-to-head comparison across rules and across methods
is therefore impossible. To the author's knowledge, no open
benchmark suite for high-entropy alloy phase classification exists
that is analogous in scope to the MatBench resource for materials
property prediction \cite{dunn2020matbench}. Second, the rules are
typically reported as accuracies of the form $X / N$, without
sensitivity, specificity, Wilson intervals, receiver-operating-
characteristic analysis, or Youden's $J$. This omission obscures
the regimes in which each rule actually generalizes. Third, the
only widely used interactive tool for Miedema-style enthalpy
calculations is the registration-gated, Windows-only MiedCalc
program of Zhang and colleagues \cite{zhang2016miedema}. No
zero-install open analogue is currently available.

The contribution of this work addresses all three. It consolidates
three primary open phase-label datasets into a single resource
with full provenance. It implements the four canonical rules as
diagnostic classifiers and reports their statistics on the
consolidated benchmark. It distributes the result in two
equivalent surfaces, namely a standard Python library and an
in-browser front end that executes the same release wheel through
the Pyodide runtime.

% ====================================================================
\section{Software description}
% ====================================================================

\subsection{Software architecture}

The package is organized as a small set of orthogonal
sub-components and a thin command-line interface. A composition
component performs formula parsing and normalization, with a
single regular-expression-based parser that accepts the three
formula conventions found in the source datasets, namely
space-separated proportional amounts, compact element-coefficient
strings, and bare formula tokens. A descriptor component
implements the six empirical descriptors required by the rule set,
namely $\Smix$, $\delta$, the valence electron concentration,
$\Tm$, $\Hmix$ \cite{miedema1973,miedema1980}, and $\Omegaparam$
\cite{yang2012}. Per-element parameters covering 24 elements are
held in an internal data table. Binary mixing-enthalpy pairs are
loaded from a vendored copy of the matminer Miedema pair table
\cite{ward2018}, licensed under BSD-3-Clause with the upstream
license file preserved alongside the data. The vendored table
derives from the de Boer parameter set \cite{deboer1988} and is
consistent with the open tabulation of Takeuchi and Inoue
\cite{takeuchi2005}.

A rules component holds the four canonical empirical rules wrapped
as classifiers, each with a tunable threshold defaulting to the
value reported in the original publication. A classifier-statistics
component provides pure-stdlib implementations of Wilson score
intervals \cite{wilson1927}, confusion matrices, binary classifier
statistics, and receiver-operating-characteristic sweeps with
Youden's $J$ \cite{youden1950}. A benchmark component holds the
per-source loaders, the cross-source consolidator, and a coverage
analysis that quantifies which fraction of the consolidated dataset
is scorable by the current descriptor data tables. An evaluation
orchestrator runs each rule against the consolidated benchmark and
writes the resulting statistics as a structured release artifact.

Two end-user surfaces consume the library. The Python surface is a
standard package-manager-installable distribution. The browser
surface is a single HTML page that loads the Pyodide runtime
\cite{pyodide} and, in turn, installs the same release wheel that
the package manager would deliver. Descriptor and rule calls then
execute client-side in the user's browser tab, with no server-side
component. The library has one canonical implementation, and both
surfaces invoke it identically.

\subsection{Software functionalities}

The package supports six primary workflows. A user can parse an
arbitrary composition formula, normalize it to mole fractions, and
compute the six standard descriptors. A user can apply each of the
four canonical empirical rules and obtain a predicted phase
classification. A user can score a user-supplied rule or
composition set against the v0.1.0 benchmark with full diagnostic
statistics. A user can sweep a rule threshold and obtain the
receiver-operating-characteristic curve, the
optimal-accuracy threshold, and the optimal Youden's-$J$ threshold.
A user can analyse element coverage of an alloy set against the
descriptor data tables and inspect a ranking of next-best-add
elements. A user can re-consolidate the per-source data through the
bundled loaders and the deduplication and conflict-detection logic.

\subsection{The consolidated benchmark}

Version 0.1.0 of the package ships a single canonical benchmark
containing \num{7784} unique compositions. It is built by running
the per-source loaders on three open datasets. The dataset of Borg
and colleagues \cite{borg2020} contributes \num{740} unique
(formula, processing) alloys, licensed under CC-BY-4.0 and
mirrored verbatim with a recorded SHA-256. The dataset of Pei and
colleagues \cite{pei2020} contributes \num{1209} unique formulas,
also CC-BY-4.0 and mirrored. The dataset of Peivaste and
colleagues \cite{peivaste2023} contributes \num{7747} unique
formulas after deduplication of an upstream-generated combinatorial
sweep. The upstream data repository for the Peivaste dataset
declares no license and is therefore not redistributed in the
package, but is fetched on user request at build time.

Compositions are merged on a four-decimal-precision composition
key. The processing column from the Borg dataset is preserved as a
side-channel rather than included in the join key, in order to keep
the canonical taxonomy composition-only. When two sources assign
different canonical labels to the same composition, the row is
flagged and the per-source labels are preserved verbatim. The
v0.1.0 release contains \num{100} such cross-source conflicts,
which represent \SI{1.3}{\percent} of the benchmark. These
conflicts are themselves a useful artifact, because they enumerate
compositions on which the open high-entropy alloy literature does
not yet agree.

The canonical phase taxonomy is deliberately coarse for v0.1.0.
The four classes are body-centred cubic, face-centred cubic,
hexagonal close packed, and multi-phase. This coarseness allows
all three sources to map into the canonical scheme without lossy
guessing. Finer per-source labels are preserved as side-channel
columns. These include the microstructure column from the Borg
dataset, which carries more than thirty sublabels, and the
twelve-category column from the Peivaste dataset, which
distinguishes intermetallic and amorphous phases.

% ====================================================================
\section{Illustrative examples}
% ====================================================================

\subsection{The Cantor alloy as a sanity check}

The canonical equimolar CoCrFeMnNi alloy of Cantor and colleagues
\cite{cantor2004} serves as the field's reference high-entropy
alloy and as a regression sanity check for the library. The six
descriptor values returned by the package for the equimolar
composition are $\Smix = R \ln 5 = \SI{13.381}{J\per(mol\cdot K)}$,
$\delta = \SI{3.164}{\percent}$, $\mathrm{VEC} = 8.000$,
$\Tm = \SI{1801.2}{K}$ (rule-of-mixtures average), $\Hmix =
\SI{-4.16}{kJ\per mol}$, and $\Omegaparam = 5.794$. All four
empirical rules classify the Cantor alloy as high-entropy class,
single-phase, and face-centred cubic, in agreement with the
original experimental characterization \cite{cantor2004}. Each of
the six values above is pinned in the regression test suite.

\subsection{Reference rule baselines on the consolidated benchmark}

Running the four rules through the evaluation orchestrator
produces the reference baselines shown in
Table~\ref{tab:rule-baselines}. The reported values are pinned by
the test suite. Any future change in the dataset, the descriptor
implementations, the vendored pair-enthalpy table, or the rule
thresholds will perturb these values and will also break the
build.

\begin{table}[ht]
\centering
\caption{Binary-classifier baselines on the v0.1.0 benchmark.}
\label{tab:rule-baselines}
\begin{tabular}{lrrrrr}
\toprule
Rule & $n$ & Accuracy & Sens. & Spec. & $J$ \\
\midrule
$\delta < \SI{6.5}{\percent}$ \cite{zhang2008} & \num{6651} & \SI{56.7}{\percent} & \SI{99.0}{\percent} & \SI{8.5}{\percent} & 0.075 \\
$\Omegaparam > 1.1$ \cite{yang2012}            & \num{6651} & \SI{54.4}{\percent} & \SI{95.8}{\percent} & \SI{7.4}{\percent} & 0.032 \\
\bottomrule
\end{tabular}
\end{table}

The Wilson 95\% confidence interval on the Zhang $\delta$ accuracy
is $[\SI{55.5}{\percent}, \SI{57.9}{\percent}]$, and on the Yang
$\Omegaparam$ accuracy is $[\SI{53.2}{\percent}, \SI{55.6}{\percent}]$.
The two binary rules return Youden's $J$ values close to zero,
which indicates that on the consolidated benchmark they collapse
to a regime of predicting single-phase almost always.

The VEC rule of Guo and Liu \cite{guo2011} predicts crystal
structure rather than single-phase formation, and is therefore
evaluated by restricting observations to those labelled
face-centred-cubic or body-centred-cubic. On the \num{3463}
single-phase observations in the benchmark, the rule attains an
accuracy of \SI{66.9}{\percent}. The per-class breakdown shows a
strong asymmetry, with \SI{92.3}{\percent} of observed
face-centred-cubic alloys predicted correctly but only
\SI{48.3}{\percent} of observed body-centred-cubic alloys.

The Yeh and colleagues \cite{yeh2004} $\Smix$ criterion is
descriptive rather than predictive. On the consolidated benchmark,
\SI{47.0}{\percent} of compositions pass the $\Smix > 1.5 R$
high-entropy threshold, \SI{36.9}{\percent} fall in the
medium-entropy range $R \leq \Smix \leq 1.5R$, and
\SI{16.1}{\percent} fall below the medium-entropy threshold.

\subsection{Threshold recalibration of the Zhang rule}

Treating the rules as diagnostic classifiers admits a question
that the canonical literature does not address. What threshold
would maximize Youden's $J$ on this benchmark? Sweeping the
Zhang \cite{zhang2008} $\delta$ threshold from \SI{1}{\percent} to
\SI{12}{\percent} in steps of \SI{0.1}{\percent} yields the result
summarized in Table~\ref{tab:roc-zhang}.

\begin{table}[h]
\centering
\caption{Receiver-operating-characteristic sweep of the Zhang
$\delta$ rule on the v0.1.0 benchmark.}
\label{tab:roc-zhang}
\begin{tabular}{lrrr}
\toprule
Threshold & Sens. & Spec. & Youden's $J$ \\
\midrule
\SI{6.5}{\percent} (canonical) & \SI{99.0}{\percent} & \SI{8.5}{\percent}  & 0.075 \\
\SI{6.6}{\percent} (max acc.)  & \SI{99.5}{\percent} & \SI{8.2}{\percent}  & 0.075 \\
\SI{2.5}{\percent} (max $J$)   & \SI{23.6}{\percent} & \SI{87.7}{\percent} & \textbf{0.113} \\
\bottomrule
\end{tabular}
\end{table}

The canonical \SI{6.5}{\percent} threshold is heavily skewed
toward sensitivity at the expense of specificity. The
balanced-optimal threshold on the consolidated benchmark is
approximately 2.6 times tighter. Whether $J = 0.113$ constitutes a
useful classifier is a separate question, and quantitatively it
does not. The recalibration finding itself is, however, a direct
and defensible output of treating the empirical rule as a
diagnostic classifier rather than as a fixed cutoff.

% ====================================================================
\section{Impact}
% ====================================================================

The contribution consolidates an open benchmark that was not
previously available in unified form, and supplies the diagnostic-
classifier statistics that were not previously applied to the
canonical high-entropy alloy rules in the open literature. Three
classes of impact are anticipated.

The first is the calibration of expectations. The collapse of both
binary rules to near-zero Youden's $J$ on the consolidated
benchmark serves as a quantitative caveat for downstream users of
those rules. The \SI{80}{\percent}-accuracy figures quoted in the
original papers reflect smaller, different datasets and a different
choice of evaluation metric. On a consolidated open dataset, the
rules behave very differently.

The second is the provision of a substrate for subsequent
methodological work. A candidate model, a CALPHAD-derived
classifier, or a new empirical descriptor can now be scored
against the same dataset and reported with the same statistical
vocabulary as the four canonical rules. The result is a
head-to-head comparison rather than incommensurable
single-publication accuracies.

The third is accessibility. The Pyodide front end allows a
researcher or student to enter a composition into a browser tab
and to obtain the same numerical outputs as a local install of the
Python package, with the release wheel itself loaded into the
browser rather than a separate JavaScript reimplementation. This
removes the local-install barrier without compromising the
library-as-canonical-implementation guarantee.

% ====================================================================
\section{Conclusions}
% ====================================================================

The article has presented \hea{} 0.0.1, an open Python library and
benchmark suite for high-entropy alloy phase prediction. The
package ships a \num{7784}-composition consolidated benchmark
drawn from three primary open sources with explicit cross-source
provenance and conflict tracking. It provides reference-baseline
diagnostic-classifier evaluations of the four canonical empirical
rules. It reports a receiver-operating-characteristic recalibration
finding for the Zhang $\delta$ rule, with a dataset-optimal
threshold of $\delta < \SI{2.5}{\percent}$ rather than the
canonical $\delta < \SI{6.5}{\percent}$. It includes an
in-browser Pyodide front end that runs the same release wheel as a
local package-manager install. The package is released under the
MIT License and is covered by more than 150 tests, including
pinned regression tests on every reported benchmark statistic. The intent
is for the artifact to serve as the substrate on which subsequent
methods for high-entropy alloy phase prediction can be compared.

Planned future work has two strands. The first is an expansion of
the elemental-property table to lift benchmark coverage from
\SI{86.7}{\percent} to approximately \SI{95}{\percent}. The
required additions are mechanical, because the corresponding
pair-enthalpy entries are already present in the vendored Miedema
table. The second is a v0.2.0 data release that exposes the
cross-source label conflicts as a structured side-dataset for
downstream disagreement-resolution work.

%=====================================================================
\section*{Conflict of interest}

The author declares no competing financial or non-financial
interests.

\section*{Acknowledgements}

The author thanks the matminer maintainers for the open Miedema
pair-enthalpy table, vendored under BSD-3-Clause, the Pyodide
project for the browser-side Python runtime, and the authors of
the three primary source datasets for releasing data under
reusable terms. This manuscript was prepared with computer-assisted
text generation. All numerical parameters, formulae, threshold
values, and benchmark results were independently verified against
the cited primary sources or computed from documented inputs by
the author.

%=====================================================================
\printbibliography

\end{document}
